\chapter*{List of publications}

\section*{International Journals}
\begin{enumerate}[
    leftmargin=*,
    label={J\arabic*.},
    ref={J\arabic*.}]
    % for citing your paper, take directly APA format from Google scholar
    % \item Surname, N., Surname, N. and Surname, N. (Year,Month). Title of the paper or journal. \emph{Place of publication}. (Vol. 123, pp. 456). EditorName.
    \item Bartalesi, V., Pratelli, N. (2024) Representing Geospatial Knowledge in Narratives.  \emph{ACM Journal on Computing and Cultural Heritage}, Just Accepted (November 2024).
    \item Bartalesi, V., Coro, G., Lenzi, E., Pratelli, N., Pagano, P., Moretti, M., Brunori, G. (2024). A Semantic Knowledge Graph of European Mountain Value Chains. \emph{Scientific Data}, 11(1), 978.
    \item Zabulis, X., Partarakis, N., Bartalesi, V., Pratelli, N., Meghini, C., Dubois, A., Moreno, I., Manitsaris, S. (2024). Multimodal Dictionaries for Traditional Craft Education. \emph{Multimodal Technologies and Interaction}, 8(7), 63.
    \item Zabulis, X., Partarakis, N., Demeridou, I., Bartalesi, V., Pratelli, N., Meghini, C., Nikolaou, N.,  Fallahian, P. (2024). Modelling and Simulation of Traditional Craft Actions. \emph{Applied Sciences}, 14(17), 7750.
    \item Bartalesi, V., Coro, G., Lenzi, E., Pagano, P., Pratelli, N. (2023). From unstructured texts to semantic story maps. \emph{International Journal of Digital Earth}, 16(1), 234–250.
    \item Bartalesi, V., Coro, G., Lenzi, E., Pratelli, N., Pagano, P., Felici, F., Moretti, M., Brunori, G. (2023). Using semantic story maps to describe a territory beyond its map. \emph{Semantic Web}, 14(6), 1255–1272.
    \item Bartalesi, V., Pratelli, N., Lenzi, E., Pontari, P. (2023). Using Semantic Web to Create and Explore an Index of Toponyms Cited in Medieval Geographical Works. \emph{ACM Journal on Computing and Cultural Heritage}, 16(2), 1–18.
    \item Zabulis, X., Partarakis, N., Demeridou, I., Doulgeraki, P., Zidianakis, E., Argyros, A., Theodoridou, M., Marketakis, Y., Meghini, C., Bartalesi, V. (2023). A roadmap for craft understanding, education, training, and preservation. \emph{Heritage}, 6(7), 5305–5328.
    \item Bartalesi, V., Metilli, D., Pratelli, N.,  Pontari, P. (2022). Towards a knowledge base of medieval and renaissance geographical Latin works: The IMAGO ontology. \emph{Digital Scholarship in the Humanities}, 37(1), 34–50.
    \item Bartalesi, V., Pratelli, N., Lenzi, E. (2022). Linking different scientific digital libraries in Digital Humanities: The IMAGO case study. \emph{International Journal on Digital Libraries}, 23(4), 303–317.
    \item Bartalesi, V., Pratelli, N., Meghini, C., Metilli, D., Tomazzoli, G., Livraghi, L. M., Zaccarello, M. (2022). A formal representation of the divine comedy’s primary sources: The Hypermedia Dante Network ontology. \emph{Digital Scholarship in the Humanities}, 37(3), 630–643.
\end{enumerate}

\section*{International Conferences/Workshops with Peer Review}
\begin{enumerate}[
    leftmargin=*,
    label={C\arabic*.},
    ref={C\arabic*.}]
    % for citing your paper, take directly APA format from Google scholar
    % \item Surname, N., Surname, N. and Surname, N. (Year,Month). Title of the paper or journal. \emph{Place of publication}. (Vol. 123, pp. 456). EditorName.
    
    \item Bartalesi, V., Lenzi, E., Pratelli, N. (2023). A Web Tool to Create and Visualise Semantic Story Maps. 163–169.
    \item Bartalesi, V., Pratelli, N. (2021). The imago project: Towards a knowledge base of medieval and renaissance geographical works. SWODCH.
    \item Bartalesi, V., Pratelli, N., Lenzi, E. (2022). A Knowledge Base of Medieval and Renaissance Geographic Latin Works. IRCDL.
    \item Pratelli, N. (2022). A Geographical Extension for NOnt Ontology. 3246, 126–133.
\end{enumerate}

\section*{Others}
\begin{enumerate}[
    leftmargin=*,
    label={O\arabic*.},
    ref={O\arabic*.}]
    % for citing your paper, take directly APA format from Google scholar
    % \item Surname, N., Surname, N. and Surname, N. (Year,Month). Title of the paper or journal. \emph{Place of publication}. (Vol. 123, pp. 456). EditorName.
    \item Bartalesi, V., Coro, G., Lenzi, E., Pratelli, N., Pagano, P. (2023). Towards Digital Twins of Territories Through Semantic Story Maps. \emph{BUILding a DIgital Twin: requirements, methods, and applications}.
    \item Bartalesi, V., Lenzi, E., Pratelli, N. (2023). Enhancing Narratives in Digital Humanities: Harnessing the Capabilities of Semantic Story Maps. Life Narrative and the Digital 2023.
\end{enumerate}














